---
title: "PopGenHelpR function testing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PopGenHelpR_viztests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  options(rmarkdown.html_vignette.check_title = FALSE)
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose
This document contains code and results from stress testing the visualization functions of PopGenHelpR (PGH). This allows us to catch any potential errors before publishing the new version of PGH to the CRAN.

**_!!! WARNING !!!_**

**_This document is not pretty, it exists to demonstrate that we have tested many different combinations of arguments in our functions and that they work as expected. Prettier documents can be found in other articles on the [PopGenHelpR website](https://kfarleigh.github.io/PopGenHelpR/)._**


## Load data and source functions
```{r load data and source functions}
library(cowplot)
library(magrittr)

# Install developmental PopGenHelpR if needed 
devtools::install_github("kfarleigh/PopGenHelpR")

.rs.restartR()

library(PopGenHelpR)


# Data

Q_dat <- PopGenHelpR::Q_dat
Pop_dat <- PopGenHelpR::HornedLizard_Pop
Fst_dat <- PopGenHelpR::Fst_dat
Het_dat <- PopGenHelpR::Het_dat


# Spatial data
shapefiles <- list.files(path = "/Users/keaka/OneDrive/Desktop/Projects/Crotalus_spatial/Final_shapefiles/",
                         pattern = "*.shp$", full.names = T)

# Remove the viridis shapefile
shapefiles <- shapefiles[1:8]

# Get elevation data 
raster <- geodata::elevation_global(path = tempdir(), res = 5)

# Get temperature data
temp_ras <- geodata::worldclim_global("tavg", path = tempdir(), res = 5)
```

### Ancestry barchart

We have updated the ancestry barchart to allow users to specify the order of individuals/populations and to handle a mix of characters and numeric information for the population and individual names. It also no longer requires the population assignment file to be four columns, just two (individual and population). The function also sorts individuals/populations by cluster if no order is specified.

```{r Ancestry barchart, out.width= 1000, out.height= 750}
# Isolate the q-matrix and population information, create Pop_mix to show it works with a mixture of character and numerics.
Qmat <- Q_dat[[1]]
Pops <- Q_dat[[2]]

Pops_mix <- Pops
Pops_mix$Sample <- as.character(Pops_mix$Sample)
```

#### **Generate ancestry matrices with mixed individual and population labels.**

First, we will show that using a mixture of character and numeric labels does not make a difference, and that PGH now sorts the ancestry matrix to cluster indviduals with similar ancestry together. We expect these plots to look exactly the same.  

```{r Ancestry barchart-1, out.width= 1000, out.height= 750 }
# Run with no order specifiied
Base <- Ancestry_barchart(anc.mat = Qmat, pops = Pops, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'))

Mix_char_num <- Ancestry_barchart(anc.mat = Qmat, pops = Pops_mix, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'))


plot_grid(Base$`Individual Ancestry Plot`, Mix_char_num$`Individual Ancestry Plot`, ncol = 1)
plot_grid(Base$`Population Ancestry Plot`, Mix_char_num$`Population Ancestry Plot`, ncol = 1)
```


#### **Generate ancestry matrices of specific plot types.**

Next, we will test the `plot.type` argument to ensure that results are consistent regardless of plot.type.

``` {r Ancestry barchart-2, out.width= 1000, out.height= 750}
# Make sure it works for only individual and population plot types 

Base_ind <- Ancestry_barchart(anc.mat = Qmat, pops = Pops, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), plot.type = "individual")

Base_ind$`Individual Ancestry Matrix`

Base_pop <- Ancestry_barchart(anc.mat = Qmat, pops = Pops, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), plot.type = "population")

Base_pop$`Population Ancestry Matrix`
```

#### **Generate ancestry matrices with specific individual and population ordering.**

Finally, we will set individual, population, and both individual and population orders for plotting to test the `ind.order` and `pop.order` arguments. We expect the individual plots to go `1-30` and for the population plots to go `1-5`.

```{r Ancestry barchart-3, out.width= 1000, out.height= 750}
# Order by individual, test when plot type is individual and all 

Base_ind_ord <- Ancestry_barchart(anc.mat = Qmat, pops = Pops, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), plot.type = "individual", ind.order = Pops$Sample)

Base_ind_ord$`Individual Ancestry Matrix`

Base_ind_ord <- Ancestry_barchart(anc.mat = Qmat, pops = Pops, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), plot.type = "all", ind.order = Pops$Sample)

Base_ind_ord$`Individual Ancestry Matrix`


# Order by population, test when plot type is population and all 

Base_pop_ord <- Ancestry_barchart(anc.mat = Qmat, pops = Pops, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), plot.type = "population", pop.order = unique(Pops$Population))

Base_pop_ord$`Population Ancestry Matrix`

Base_pop_ord <- Ancestry_barchart(anc.mat = Qmat, pops = Pops, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), plot.type = "all", pop.order = unique(Pops$Population))

Base_pop_ord$`Population Ancestry Matrix`

# Specify both an individual and population order, this scenario is only relevant for plot type all

Base_ind_pop_ord <- Ancestry_barchart(anc.mat = Qmat, pops = Pops, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), plot.type = "all", pop.order = unique(Pops$Population), ind.order = Pops$Sample)

# Individual plot 
Base_ind_pop_ord$`Individual Ancestry Plot`

# Population plot
Base_ind_pop_ord$`Population Ancestry Plot`
```

Let's move onto the `Piechart_map` function.

### Piechart map

Here, the new functionality is the ability to add shapefiles, a scale bar, and a north arrow to the map. Users can change the position of the shapefile, to be under boundaries (i.e., state lines), ontop of boundaries, and ontop of everything. **Note that the shapefiles do not represent the ranges of the species piecharts that are being plotted; this is only to demonstrate functionality.** Also, if the shapefile is on top of everything else, it will cause the coordinate system to expand (plot will be sized differently).

#### **Generate original PGH piechart maps**

First, we will generate piechart maps without extra layers, a north arrow, or scale bar. We show that PGH works with differnt plot types. 

```{r Piechart map, out.width= 1000, out.height= 1250}

# Make map without shapefiles, this is what the original version of PGH did.
Base_piemap <- Piechart_map(anc.mat = Qmat, pops = Pops, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), plot.type = "all", Lat_buffer = 3, Long_buffer = 3, country_code = c("usa", "can", "mex"))

plot_grid(Base_piemap$Individual_piemap, Base_piemap$Population_piemap, nrow = 1)

# Show it works for individual and population plot.types
Base_piemap <- Piechart_map(anc.mat = Qmat, pops = Pops, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), plot.type = "individual", Lat_buffer = 3, Long_buffer = 3, country_code = c("usa", "can", "mex"))

Base_piemap$Individual_piemap

Base_piemap <- Piechart_map(anc.mat = Qmat, pops = Pops, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), plot.type = "population", Lat_buffer = 3, Long_buffer = 3, country_code = c("usa", "can", "mex"))

Base_piemap$Population_piemap
```

#### **Generate PGH piechart maps with shapefiles, different shapefile positions**

Now, we will add shapefiles to the plots and demonstrate how the `shapefile_plot_position` argument works. We also demonstrate how to plot transparent shapefiles that are outlined. **It is expected that you will receive a warning that the scale on the map varies by more than 10%, scale bar may be inaccurate because the data are not in a projected coordinate system**. We do not support projection of data within PGH because we do not want to decide which projection is best for your project and because projecting large shapefiles and rasters can crash R. 

```{r Piechart map-1, out.width= 1000, out.height= 1250}
# Try different shapefile positions
Shap_piemap <- Piechart_map(anc.mat = Qmat, pops = Pops, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), plot.type = "all", Lat_buffer = 3, Long_buffer = 3, country_code = c("usa", "can", "mex"), shapefile = shapefiles, shapefile_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'), shapefile_plot_position = 1)
 
Shap1_piemap <- Piechart_map(anc.mat = Qmat, pops = Pops, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), plot.type = "all", Lat_buffer = 3, Long_buffer = 3, country_code = c("usa", "can", "mex"), shapefile = shapefiles, shapefile_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'), shapefile_plot_position = 1, north_arrow = T, scale_bar = T, north_arrow_position = "tr") 

Shap2_piemap <- Piechart_map(anc.mat = Qmat, pops = Pops, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), plot.type = "all", Lat_buffer = 3, Long_buffer = 3, country_code = c("usa", "can", "mex"), shapefile = shapefiles, shapefile_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'), shapefile_plot_position = 2,north_arrow = T, scale_bar = T, north_arrow_position = "tr") 

Shap3_piemap <- Piechart_map(anc.mat = Qmat, pops = Pops, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), plot.type = "all", Lat_buffer = 3, Long_buffer = 3, country_code = c("usa", "can", "mex"), shapefile = shapefiles, shapefile_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'), shapefile_plot_position = 3,north_arrow = T, scale_bar = T, north_arrow_position = "tr") 

plot_grid(Shap1_piemap$Population_piemap, Shap2_piemap$Population_piemap, Shap3_piemap$Population_piemap, nrow = 1)

plot_grid(Shap1_piemap$Individual_piemap, Shap2_piemap$Individual_piemap, Shap3_piemap$Individual_piemap, nrow = 1)


# Make a plot with a transparent shapefile 

Shapoutlined_piemap <- Piechart_map(anc.mat = Qmat, pops = Pops, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), plot.type = "all", Lat_buffer = 3, Long_buffer = 3, country_code = c("usa", "can", "mex"), shapefile = shapefiles, shapefile_outline_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'), shapefile_col = rep(NA,8), shapefile_plot_position = 2,north_arrow = T, scale_bar = T, north_arrow_position = "tr", shp_outwidth = 2) 

plot_grid(Shapoutlined_piemap$Individual_piemap, Shapoutlined_piemap$Population_piemap, nrow = 1)
```


The `Piechart_map` function also seems to work. Let's move onto the `Plot_coordinates` function.

### Plot coordinates
Users can now add shapefiles and rasters to the maps output by `Plot_coordinates`, along with north arrows and scale bars. Now, they can also change the colors of the points based on group assignment.

First, we will generate a plot with points and no raster or shapefile as a base layer. We can generate a map where poitns are all the same color (no `group` or `group_col` argument) or a map where points are colored by specifying the `group` or `group_col` arguments.

```{r Plot Coordinates, out.width= 1000, out.height= 1250}

Base_PC_nogrp <- Plot_coordinates(dat = Pop_dat, Lat_buffer = 3, Long_buffer = 3, Longitude_col = 3, Latitude_col = 3, country_code = c("usa", "mex", "can"))

Base_PC_wgrp <- Plot_coordinates(dat = Pop_dat, group = Pop_dat$Population,  group_col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), Lat_buffer = 3, Long_buffer = 3, Longitude_col = 3, Latitude_col = 3, country_code = c("usa", "mex", "can"))

plot_grid(Base_PC_nogrp, Base_PC_wgrp, nrow = 1)

```

Now, let's add a shapefile, scale bar, and north arrow.

```{r Plot Coordinates-1, out.width= 1000, out.height= 1250}

PC_shp1 <- Plot_coordinates(dat = Pop_dat, Lat_buffer = 3, Long_buffer = 3, Longitude_col = 3, Latitude_col = 4, group = Pop_dat$Population,  group_col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), country_code = c("usa", "mex", "can"), shapefile = shapefiles, shapefile_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'), shapefile_plot_position = 1,north_arrow = T, scale_bar = T, north_arrow_position = "tr", shapefile_outline_col = NA)

PC_shp2 <- Plot_coordinates(dat = Pop_dat, Lat_buffer = 3, Long_buffer = 3, Longitude_col = 3, Latitude_col = 4, group = Pop_dat$Population,  group_col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), country_code = c("usa", "mex", "can"), shapefile = shapefiles, shapefile_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'), shapefile_plot_position = 2,north_arrow = T, scale_bar = T, north_arrow_position = "tr", shapefile_outline_col = NA)

PC_shp3 <- Plot_coordinates(dat = Pop_dat, Lat_buffer = 3, Long_buffer = 3, Longitude_col = 3, Latitude_col = 4, group = Pop_dat$Population,  group_col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), country_code = c("usa", "mex", "can"), shapefile = shapefiles, shapefile_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'), shapefile_plot_position = 3,north_arrow = T, scale_bar = T, north_arrow_position = "tr", shapefile_outline_col = NA)

PC_shpout <- Plot_coordinates(dat = Pop_dat, Lat_buffer = 3, Long_buffer = 3, Longitude_col = 3, Latitude_col = 4, group = Pop_dat$Population,  group_col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), country_code = c("usa", "mex", "can"), shapefile = shapefiles, shapefile_outline_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'), shapefile_col = rep(NA,8), shapefile_plot_position = 3,north_arrow = T, scale_bar = T, north_arrow_position = "tr")


plot_grid(PC_shp1, PC_shp2, PC_shp3, PC_shpout, nrow = 2, ncol = 2)

```


Now, let's add a raster. **Sometimes `terra` interprets rasters as discrete, even though they are not. We have added the `disrete_raster` argument to accomodate this, but fair warning that it can throw an error the first time you try your raster**. The solution is just to change the `discrete_raster` argument to TRUE or FALSE (depending on what failed).

```{r Plot Coordinates-2, out.width= 1000, out.height= 1250}

PC_ras1 <- Plot_coordinates(Pop_dat, Longitude_col = 3, Latitude_col = 4, group = Pop_dat$Population, group_col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), country_code = c("usa", "mex", "can"), raster = raster, raster_plot_position = 1, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = TRUE,    
raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), raster_breaks = c(0,500,1000,1500,2000,2500,3000,3500,4000,5000), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE)

PC_ras2 <- Plot_coordinates(Pop_dat, Longitude_col = 3, Latitude_col = 4, group = Pop_dat$Population, group_col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), country_code = c("usa", "mex", "can"), raster = raster, raster_plot_position = 2, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = TRUE,    
raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), raster_breaks = c(0,500,1000,1500,2000,2500,3000,3500,4000,5000), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE)

PC_ras3 <- Plot_coordinates(Pop_dat, Longitude_col = 3, Latitude_col = 4, group = Pop_dat$Population, group_col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), country_code = c("usa", "mex", "can"), raster = raster, raster_plot_position = 3, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = TRUE,    
raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), raster_breaks = c(0,500,1000,1500,2000,2500,3000,3500,4000,5000), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE)

plot_grid(PC_ras1, PC_ras2, PC_ras3)


PC_temp <-  Plot_coordinates(Pop_dat, Longitude_col = 3, Latitude_col = 4, group = Pop_dat$Population, group_col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), country_code = c("usa", "mex", "can"), raster = raster, raster_plot_position = 3, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = FALSE,    
raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE)

PC_temp_wbreaks <- Plot_coordinates(Pop_dat, Longitude_col = 3, Latitude_col = 4, group = Pop_dat$Population, group_col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), country_code = c("usa", "mex", "can"), raster = temp_ras, raster_plot_position = 1, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = FALSE,    
raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE, raster_breaks = c(-15, -10, -5,-2.5, 0,2.5, 5, 10, 15, 20))

plot_grid(PC_temp, PC_temp_wbreaks)

```

Finally, we will create the map with a raster and shapefiles. 

```{r Plot Coordinates-3, out.width= 1000, out.height= 1250}

PC_ras_shp <- Plot_coordinates(Pop_dat, Longitude_col = 3, Latitude_col = 4, group = Pop_dat$Population, group_col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'), country_code = c("usa", "mex", "can"), shapefile = shapefiles, shapefile_plot_position = 1, raster = raster, raster_plot_position = 2, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = TRUE, shapefile_outline_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'),    raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), raster_breaks = c(0,500,1000,1500,2000,2500,3000,3500,4000,5000), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE)

PC_ras_shp

```

### Point map 
Users can now add shapefiles and rasters to the maps output by `Point_map`, along with north arrows and scale bars.  

First, we will just make a `Point_map` map with no background, just the regular basemap and administrative boundaries.

```{r Point map-1, out.width= 1000, out.height= 1250}

Base_PM <- Point_map(dat = Het_dat, statistic = "Heterozygosity", country_code = c("usa", "can", "mex"))

# Set an outline color
Base_PM_out <- Point_map(dat = Het_dat, statistic = "Heterozygosity", country_code = c("usa", "can", "mex"), out.col = "black")

plot_grid(Base_PM$`Heterozygosity Map`, Base_PM_out$`Heterozygosity Map`, nrow = 1)

```


Now let's add shapefiles, a scale bar, and north arrow.

```{r Point map-2, out.width= 1000, out.height= 1250}

PM_shp1 <- Point_map(dat = Het_dat, statistic = "Heterozygosity", Lat_buffer = 3, Long_buffer = 3, Longitude_col = 4, Latitude_col = 5, country_code = c("usa", "mex", "can"), shapefile = shapefiles, shapefile_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'), shapefile_plot_position = 1,north_arrow = T, scale_bar = T, north_arrow_position = "tr", shapefile_outline_col = NA, out.col = "black")

PM_shp2 <- Point_map(dat = Het_dat, statistic = "Heterozygosity", Lat_buffer = 3, Long_buffer = 3, Longitude_col = 4, Latitude_col = 5, country_code = c("usa", "mex", "can"), shapefile = shapefiles, shapefile_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'), shapefile_plot_position = 2,north_arrow = T, scale_bar = T, north_arrow_position = "tr", shapefile_outline_col = NA, out.col = "black")

PM_shp3 <- Point_map(dat = Het_dat, statistic = "Heterozygosity", Lat_buffer = 3, Long_buffer = 3, Longitude_col = 4, Latitude_col = 5, country_code = c("usa", "mex", "can"), shapefile = shapefiles, shapefile_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'), shapefile_plot_position = 3,north_arrow = T, scale_bar = T, north_arrow_position = "tr", shapefile_outline_col = NA, out.col = "black")

PM_shpout <- Point_map(dat = Het_dat, statistic = "Heterozygosity", Lat_buffer = 3, Long_buffer = 3, Longitude_col = 4, Latitude_col = 5, country_code = c("usa", "mex", "can"), shapefile = shapefiles, shapefile_outline_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'), shapefile_col = rep(NA,8), shapefile_plot_position = 3,north_arrow = T, scale_bar = T, north_arrow_position = "tr", out.col = "black")


plot_grid(PM_shp1$`Heterozygosity Map`, PM_shp2$`Heterozygosity Map`, PM_shp3$`Heterozygosity Map`, PM_shpout$`Heterozygosity Map`)

```


Next, let's add a raster. Remember that funky things can happen with `terra` interpreting your raster as discrete or continous. Use the `discrete_raster` arugment to get around this.

```{r Point map-3, out.width= 1000, out.height= 1250}
PM_ras1 <- Point_map(Het_dat, statistic = "Heterozygosity", Longitude_col = 4, Latitude_col = 5, country_code = c("usa", "mex", "can"), raster = raster, raster_plot_position = 1, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = TRUE,    
raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), raster_breaks = c(0,500,1000,1500,2000,2500,3000,3500,4000,5000), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE)

PM_ras2 <- Point_map(Het_dat,  statistic = "Heterozygosity", Longitude_col = 4, Latitude_col = 5, country_code = c("usa", "mex", "can"), raster = raster, raster_plot_position = 2, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = TRUE,    
raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), raster_breaks = c(0,500,1000,1500,2000,2500,3000,3500,4000,5000), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE)

PM_ras3 <- Point_map(Het_dat, statistic = "Heterozygosity",  Longitude_col = 4, Latitude_col = 5, country_code = c("usa", "mex", "can"), raster = raster, raster_plot_position = 3, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = TRUE,    
raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), raster_breaks = c(0,500,1000,1500,2000,2500,3000,3500,4000,5000), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE)

plot_grid(PM_ras1$`Heterozygosity Map`, PM_ras2$`Heterozygosity Map`, PM_ras3$`Heterozygosity Map`)


PM_temp <-  Point_map(Het_dat, statistic = "Heterozygosity",  Longitude_col = 4, Latitude_col = 5, country_code = c("usa", "mex", "can"), raster = raster, raster_plot_position = 3, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = FALSE,    
raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE)

PM_temp_wbreaks <- Point_map(Het_dat, statistic = "Heterozygosity", Longitude_col = 4, Latitude_col = 5, country_code = c("usa", "mex", "can"), raster = temp_ras, raster_plot_position = 1, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = FALSE,    
raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE, raster_breaks = c(-15, -10, -5,-2.5, 0,2.5, 5, 10, 15, 20))

plot_grid(PM_temp$`Heterozygosity Map`, PM_temp_wbreaks$`Heterozygosity Map`)
```

Finally, add shapefiles and a raster. 

```{r Point map-4, out.width= 1000, out.height= 1250}

PM_ras_shp <- Point_map(Het_dat, statistic = "Heterozygosity", Longitude_col = 4, Latitude_col = 5, country_code = c("usa", "mex", "can"), shapefile = shapefiles, shapefile_plot_position = 1, raster = raster, raster_plot_position = 2, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = TRUE, shapefile_outline_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'),    raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), raster_breaks = c(0,500,1000,1500,2000,2500,3000,3500,4000,5000), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE, shp_outwidth = 2)

PM_ras_shp$`Heterozygosity Map`

```

### Network map 
Users can now add shapefiles and rasters to the maps output by `Network_map`, along with north arrows and scale bars. This function has also been updated to fix a bug that removed sites if they had a statistic of 0 (i.e, Fst = 0). Originally, this was to remove localities mapping with themselves; now we remove these instances using matching names. 

First, let's create a basic map with just the edges and point.

```{r Network map-1, out.width= 1000, out.height= 1250}

Base_NM <- Network_map(Fst_dat[[1]], pops = Fst_dat[[2]], neighbors = 2, statistic = "Fst", Lat_buffer = 3, Long_buffer = 3,country_code = c("usa", "mex", "can"))

Base_NM$Map

# Specify a comparison using population names

Base_NM_comp <- Network_map(Fst_dat[[1]], pops = Fst_dat[[2]], neighbors = "South_West", statistic = "Fst", Lat_buffer = 3, Long_buffer = 3,country_code = c("usa", "mex", "can"))

plot_grid(Base_NM$Map, Base_NM_comp$Map)
```

Now, let's add a shapefile, scale bar, and north arrow.

```{r Network map-2, out.width= 1000, out.height= 1250}

NM_1 <- Network_map(Fst_dat[[1]], pops = Fst_dat[[2]], neighbors = 2, statistic = "Fst", Lat_buffer = 3, Long_buffer = 3,country_code = c("usa", "mex", "can"), shapefile = shapefiles, shapefile_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'), shapefile_plot_position = 1,north_arrow = T, scale_bar = T, north_arrow_position = "tr", shapefile_outline_col = NA)

NM_2 <- Network_map(Fst_dat[[1]], pops = Fst_dat[[2]], neighbors = 2, statistic = "Fst", Lat_buffer = 3, Long_buffer = 3,country_code = c("usa", "mex", "can"), shapefile = shapefiles, shapefile_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'), shapefile_plot_position = 2,north_arrow = T, scale_bar = T, north_arrow_position = "tr", shapefile_outline_col = NA)

NM_3 <- Network_map(Fst_dat[[1]], pops = Fst_dat[[2]], neighbors = 2, statistic = "Fst", Lat_buffer = 3, Long_buffer = 3,country_code = c("usa", "mex", "can"), shapefile = shapefiles, shapefile_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'), shapefile_plot_position = 3,north_arrow = T, scale_bar = T, north_arrow_position = "tr", shapefile_outline_col = NA)

NM_4 <- Network_map(Fst_dat[[1]], pops = Fst_dat[[2]], neighbors = 2, statistic = "Fst", Lat_buffer = 3, Long_buffer = 3,country_code = c("usa", "mex", "can"), shapefile = shapefiles, shapefile_outline_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'), shapefile_plot_position = 1,north_arrow = T, scale_bar = T, north_arrow_position = "tr", shp_outwidth = 2)

plot_grid(NM_1$Map, NM_2$Map, NM_3$Map, NM_4$Map)

```

Next, let's add a raster.

```{r Network map-3, out.width= 1000, out.height= 1250}

NM_ras1 <- Network_map(Fst_dat[[1]], pops = Fst_dat[[2]], statistic = "Fst", neighbors = 2, country_code = c("usa", "mex", "can"), raster = raster, raster_plot_position = 1, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = TRUE,    
raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), raster_breaks = c(0,500,1000,1500,2000,2500,3000,3500,4000,5000), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE)

NM_ras2 <- Network_map(Fst_dat[[1]], pops = Fst_dat[[2]], statistic = "Fst",neighbors = 2, country_code = c("usa", "mex", "can"), raster = raster, raster_plot_position = 2, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = TRUE,    
raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), raster_breaks = c(0,500,1000,1500,2000,2500,3000,3500,4000,5000), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE)

NM_ras3 <- Network_map(Fst_dat[[1]], pops = Fst_dat[[2]], statistic = "Fst",  neighbors = 2, country_code = c("usa", "mex", "can"), raster = raster, raster_plot_position = 3, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = TRUE,    
raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), raster_breaks = c(0,500,1000,1500,2000,2500,3000,3500,4000,5000), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE)

plot_grid(NM_ras1$Map, NM_ras2$Map, NM_ras3$Map)


NM_temp <-  Network_map(Fst_dat[[1]], pops = Fst_dat[[2]],  statistic = "Fst",  neighbors = 2, country_code = c("usa", "mex", "can"), raster = raster, raster_plot_position = 1, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = FALSE,    
raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE)

NM_temp_wbreaks <- Network_map(Fst_dat[[1]], pops = Fst_dat[[2]], statistic = "Fst", neighbors = 2, country_code = c("usa", "mex", "can"), raster = temp_ras, raster_plot_position = 1, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = FALSE,    
raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE, raster_breaks = c(-15, -10, -5,-2.5, 0,2.5, 5, 10, 15, 20))

plot_grid(NM_temp$Map, NM_temp_wbreaks$Map)
```


Finally, let's make a map with both a shapefile and raster. 

```{r Network map-4, out.width= 1000, out.height= 1250}
NM_ras_shp <- Network_map(dat = Fst_dat[[1]], pops = Fst_dat[[2]],  statistic = "Fst",  neighbors = 2, country_code = c("usa", "mex", "can"), shapefile = shapefiles, shapefile_plot_position = 1, raster = raster, raster_plot_position = 2, interpolate_raster = TRUE, Lat_buffer = 3, Long_buffer = 3, discrete_raster = TRUE, shapefile_outline_col = c('#8c510a','#d8b365','#f6e8c3','#f5f5f5','#c7eae5','#5ab4ac','#01665e'),    raster_col = c('white','#ffffcc','#ffeda0','#fed976','#feb24c','#fd8d3c','#fc4e2a','#e31a1c','#bd0026','#800026'), raster_breaks = c(0,500,1000,1500,2000,2500,3000,3500,4000,5000), legend_pos = "right", scale_bar = TRUE, north_arrow_position = "tr", north_arrow = TRUE, shp_outwidth = 2)

NM_ras_shp$Map

```



### Pairwise heatmap
Users now have more freedom to control the color gradient.  

```{r Pairwise heatmap , out.width= 1000, out.height= 1250}

Base_PH <- Pairwise_heatmap(Fst_dat[[1]], statistic = "Fst")

PH_wbreaks <- Pairwise_heatmap(Fst_dat[[1]], statistic = "Fst", breaks = c(0.2,0.225,0.25,0.3,0.38))

plot_grid(Base_PH, PH_wbreaks)

```



**Thank you for your interest in our package and for ready through all of this. Please reach out if you have any questions or suggestions that could help make the package better.**
