---
title: "Generating confidence intervals with bootstrapping"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PopGenHelpR_Uncertainty}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  options(rmarkdown.html_vignette.check_title = FALSE)
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Purpose

To estimate uncertainty (confidence intervals) by performing bootstrapping.

This vignette will show users how to estimate confidence intervals using PopGenHelpR since the functions do not currently support bootstrapping.

### Why should you estimate confidence intervals?

Estimating confidence intervals provides an idea of uncertainty about a particular measurement. This is useful, because it indicates how precise our estimate is. Larger intervals (wider) indicates less precision and smaller (narrower) intervals indicates greater precision.

### What does a confidence interval indicate and what does it not indicate?

A confidence interval does *NOT* mean that there is an X% chance that your true population value lies within your interval. For example, you CANNOT say that there is a 95% chance that your true mean of a population lies within your confidence interval.

What you *CAN* say if you use a 95% confidence interval, for example, is that if you calculated this interval many times, the true mean of the population would be in 95% of those intervals.

### Notes

We will use the data that come with PopGenHelpR. There are only 3 populations, so we are able to manually extract data in our for loops below (e.g., east_south <- c(east_south, dif_boot$Fst[2,1])). This may not be feasible if you have many populations; please reach out to Keaka Farleigh (PopGenHelpR maintainer) if you are concerned about doing this manually.

### Bootstrapping for differentiation

First, we will estimate differentiation using all of the SNPs in the dataset. Then, we will perform bootstrapping by building a for loop to generate a 95% confidence interval.

```{r fst, echo=TRUE, eval=FALSE}

# Load package and data 
library(PopGenHelpR)


data("HornedLizard_Pop")
data("HornedLizard_VCF")

# set a seed to make it reproducible (I usually choose the date
set.seed(02172026)


# Estimate differentiation using all SNPs
emp_fst <- Differentiation(HornedLizard_VCF, HornedLizard_Pop, statistic = "Fst")

# Set the number of bootstrap iterations
boots <- 1000

# Get the number of SNPs
total_snps <- nrow(HornedLizard_VCF@fix)

# Create vectors to store the bootstrapped estimates for each population pair
east_west <- c()
east_south <- c()
south_west <- c()

for(i in 1:boots){
  
  loci_sample <- sample(1:total_snps,
                        size = total_snps,
                        replace = TRUE)
  
  vcf_boot <- HornedLizard_VCF[loci_sample, ]
  
  # Replace SNP names to avoid an error
  vcf_boot@fix[,"ID"] <- paste0("SNP_", 1:nrow(vcf_boot@fix))
  
  dif_boot <- suppressMessages(Differentiation(vcf_boot, HornedLizard_Pop, statistic = "Fst"))
  
  east_south <- c(east_south, dif_boot$Fst[2,1])
  east_west <- c(east_west, dif_boot$Fst[3,1])
  south_west <- c(south_west, dif_boot$Fst[3,2])
  
  remove(loci_sample, vcf_boot, dif_boot)
  
}

# Get the 95% confidence interval for each comparison
es_ci <- quantile(east_south, c(0.025, 0.975), na.rm = TRUE)
ew_ci <- quantile(east_west, c(0.025, 0.975), na.rm = TRUE)
sw_ci <- quantile(south_west, c(0.025, 0.975), na.rm = TRUE)

# Print out the confidence intervals
es_ci
ew_ci
sw_ci

# Look at the empirical values
emp_fst$Fst
```

Now we have our genome-wide estimate, and a 95% confidence interval. Generally, we expect our genome-wide estimate to fall in the middle of this confidence interval. There are exceptions, however, such as when you use very few bootstrap iterations. I recommend using at least 1,000 if you can.

Next, we will do the same thing for heterozygosity.

### Bootstrapping for heterozygosity

```{r het, echo=TRUE, eval=FALSE}

emp_obshet <- Heterozygosity(HornedLizard_VCF, HornedLizard_Pop, statistic = "Ho")


# Set the number of bootstrap replicates
boots <- 1000

# Get the number of SNPs
total_snps <- nrow(HornedLizard_VCF@fix)

# Create vectors to store the bootstrapped estimates for each population pair
east_west_obshet <- c()
east_south_obshet <- c()
south_west_obshet <- c()

for(i in 1:boots){
  
  loci_sample <- sample(1:total_snps,
                        size = total_snps,
                        replace = TRUE)
  
  vcf_boot <- HornedLizard_VCF[loci_sample, ]
  
  # Replace SNP names to avoid an error (we can't have SNPs with the same name)
  vcf_boot@fix[,"ID"] <- paste0("SNP_", 1:nrow(vcf_boot@fix))
  
  het_boot <- suppressMessages(Heterozygosity(vcf_boot, HornedLizard_Pop, statistic = "Ho"))
  
  east_south_obshet <- c(east_south_obshet, het_boot$Ho_perpop[1,1])
  east_west_obshet <- c(east_west_obshet, het_boot$Ho_perpop[2,1])
  south_west_obshet <- c(south_west_obshet, het_boot$Ho_perpop[3,1])
  
  remove(loci_sample, vcf_boot, het_boot)
  
}


# Get the 95% confidence interval for each comparison
es_obshet_ci <- quantile(east_south_obshet, c(0.025, 0.975), na.rm = TRUE)
ew_obshet_ci <- quantile(east_west_obshet, c(0.025, 0.975), na.rm = TRUE)
sw_obshet_ci <- quantile(south_west_obshet, c(0.025, 0.975), na.rm = TRUE)

# Print out the confidence intervals
es_obshet_ci
ew_obshet_ci
sw_obshet_ci

# Look at our empirical values
emp_obshet$Ho_perpop

```

Again, we should see that our genome-wide (empirical) estimates sit roughly in the middle of our confidence interval.

### Questions???

Please email Keaka Farleigh (keakafarleigh@virginia.edu) if you need help performing bootstrapping or with anything else.
