---
title: "Formatting PopGenHelpR plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PopGenHelpR_plotformatting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  options(rmarkdown.html_vignette.check_title = FALSE)
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Purpose
This document contains examples of how to format `PopGenHelpR` plots after they are generated, but still within R.

## How to format plots after they are made?

`PopGenHelpR` plots are created using ggplot2, which means that we can modify them after the fact. Normally, this means modifying the `theme`, where we can change font size, axis titles, legends, and much more (see https://ggplot2.tidyverse.org/reference/theme.html for ideas). We will go through a couple of examples to get you started. Please reach out if you have any questions or need help.

## Load the data and packages
```{r Load data}
# Load PopGenHelpR
library(PopGenHelpR)
library(cowplot)
library(ggplot2)

# Load data
data("Q_dat")

# First, we separate the list elements into two separate objects. The q-matrix (Qmat) and the locality information for each individual (Loc).
Qmat <- Q_dat[[1]]
Pops <- Q_dat[[2]]
```


We will do this using the `Ancestry_barchart` output as an example, but this can be done with any of the plots that `PopGenHelpR` produces. 

```{r Ancestry, out.width= 1500, out.height = 1250}

Base <- Ancestry_barchart(anc.mat = Qmat, pops = Pops, K = 5, col = c('#d73027', '#fc8d59', '#e0f3f8', '#91bfdb', '#4575b4'))

No_legend <- Base$`Individual Ancestry Plot` + theme(legend.position = "none")

No_legend_lab <- Base$`Individual Ancestry Plot` + theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust = 0.5))

plot_grid(Base$`Individual Ancestry Plot`, No_legend, No_legend_lab, ncol = 1)

```
