---
title: "Creating a Q-matrix to use in PopGenHelpR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PopGenHelpR_createQmatrix}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  options(rmarkdown.html_vignette.check_title = FALSE)
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Purpose

To generate a q-matrix of ancestry coefficients for use in the `PopGenHelpR` functions `Ancestry_barchart` and `Piechart_map`.

## What is a Q-matrix?

A q-matrix is a matrix containing as many rows as individuals and columns as genetic clusters. Each cell represents an ancestry coefficient (also known as cluster assignments), which is the contribution of a genetic cluster to a particular individual. Q-matrices are commonly used in population genomics to evaluate gene flow between populations (e.g., admixture) or species (e.g., introgression). ADMIXTURE (Alexander et al., 2009) and sNMF (Frichot et al., 2014) are commonly used software to estimate the number of genetic clusters in data and generate ancestry bar charts with q-matrices. 

Let's generate some q-matrices now that you know what they are! We will create q-matrices using each of the programs mentioned above. 

### sNMF

We will start with sNMF because it is implemented in the R package [LEA](https://bioconductor.org/packages/release/bioc/html/LEA.html) (Frichot & Francois, 2015). After running sNMF (see [this tutorial](https://bioconductor.org/packages/release/bioc/vignettes/LEA/inst/doc/LEA.pdf) if you need help) you just need to use the `Q` function. 

```{r sNMF, echo=TRUE, eval=FALSE}
# If I have a sNMF project named sNMFobject with K number of ancestral populations (genetic clusters), and my best run is run 1 (determined as the run with the lowets cross-entropy)
Qmat <- Q(sNMFobject, K = K, run = 1)
```

All you need to do now is append the sample names to the q-matrix as the first column (you can do this with `cbind` or in any text editor). Then you can use it in `PopGenHelpR`. Note that you must be careful that your order in the q-matrix is the same as the order of the samples you are appending. 

### ADMIXTURE

ADMIXTURE is a little more complex because it is not associated with an R package, but it is nice because it gives us the q-matrix automatically. See [this tutorial](https://speciationgenomics.github.io/ADMIXTURE/) for more details. 

In the example below, we tell ADMIXTURE to use a bed file as input and run the analysis with a K value of 5. 

```{r echo=TRUE, eval=FALSE}
### Run ADMIXTURE
admixture --cv my_genetic_data.bed 5 > K5.out
```

This will output a file with the .Q extension, which contains ancestry coefficients for each individual (our q-matrix). 

## Questions???

Please email Keaka Farleigh (farleik@miamioh.edu) if you need help generating a q-matrix or with anything else. 

## References 

Alexander, D. H., Novembre, J., & Lange, K. (2009). Fast model-based estimation of ancestry in unrelated individuals. Genome research, 19(9),     1655-1664. 

Frichot, E., & François, O. (2015). LEA: An R package for landscape and ecological association studies. Methods in Ecology and Evolution,        6(8), 925-929.

Frichot, E., Mathieu, F., Trouillon, T., Bouchard, G., & François, O. (2014). Fast and efficient estimation of individual ancestry               coefficients. Genetics, 196(4), 973-983.


